---
title: "vaccine-distribution"
author: "Anthony Morada"
date: "1/13/2021"
output:
  html_document: 
    toc: yes
    number_sections: yes
    df_print: default
  pdf_document: 
    toc: yes
    number_sections: yes
    fig_caption: yes
  word_document: 
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---
# Getting and Cleaning Data
```{r SETUP, echo=TRUE, message=TRUE, warning=TRUE, paged.print=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)

library(dplyr)
library(janitor)
library(tableone)
library(finalfit)
library(ggplot2)

#For TableOne
kableone <- function(x, ...) {
  capture.output(x <- print(x))
  knitr::kable(x, ...)
}

#Chunk timing
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      # record the current time before each chunk
      now <<- Sys.time()
    } else {
      # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now, units = "hours")
      # return a character string to show the time
      paste("Time for the chunk", options$label, "to run:", res, "hours")
    }
  }
}))

knitr::opts_chunk$set(time_it = TRUE)
```

```{r LOAD DATA}
df <- read.csv("~/Dropbox/My Mac (MacBook-Pro.hsd1.pa.comcast.net)/Documents/GitHub/covid_case_restricted_detailed/data/COVID_Cases_Restricted_Detailed_12312020_Part_1.csv")

temp <- read.csv("~/Dropbox/My Mac (MacBook-Pro.hsd1.pa.comcast.net)/Documents/GitHub/covid_case_restricted_detailed/data/COVID_Cases_Restricted_Detailed_12312020_Part_2.csv/COVID_Cases_Restricted_Detailed_12312020_Part_2.csv", header=FALSE)

colnames(temp) <- colnames(df)

df <- rbind(df,temp)

rm(temp)
```

```{r CLEAN}
# Replacing "Missing"
df<- as.data.frame(lapply(df, function(y) gsub("Missing", NA, y)))

#Formatting date columns
df <- df %>%
  mutate(cdc_report_dt=as.Date.character(cdc_report_dt, format="%Y-%m-%d")) %>%
  mutate(onset_dt=as.Date.character(onset_dt, format="%Y-%m-%d")) %>%
  mutate(pos_spec_dt=as.Date.character(pos_spec_dt, format="%Y-%m-%d")) %>%
  mutate(cdc_case_earliest_dt=as.Date.character(cdc_case_earliest_dt, format="%Y-%m-%d")) 

#Selecting for only laboratory confirmed cases
df <- df[df$current_status=="Laboratory-confirmed case",]

#Selecting for positive specimen date < 2020-12-15
df <- df[df$pos_spec_dt < "2020-12-15" & !is.na(df$pos_spec_dt),]

#Cleaning sex
df$sex <- as.factor(df$sex)
levels(df$sex) <- list(
  "Male" = "Male",
  "Female" = "Female",
  "Other or Unknown" = c("Other", "Unknown")
)

#Creating Region Variable
df$res_region <- as.factor(df$res_state)
levels(df$res_region) <- list(
  "Northeast" = c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "NY", "NYC", "PA"),
  "Midwest" = c("IN", "IL", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND","SD"),
  "South" = c("DE", "DC", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX"),
  "Pacific"= c("AZ", "CO", "ID", "NM", "MT", "UT", "NV", "WY", "AK", "CA", "HI", "OR", "WA"),
  "US Territory or Unknown" = c("GU", "MP", "OCONUS", "PR", "Unknown", "VI")
  )

```

# Exploratory Data Analysis  

```{r CATEGORIZING VARIABLES}
# independent variables
ind_vars <- c("sex", "age_group", "race_ethnicity_combined", "res_region", "hc_work_yn", "medcond_yn")
# dependent variables
dep_vars <- c("hosp_yn", "icu_yn", "mechvent_yn", "pna_yn", "acuterespdistress_yn", "death_yn")

# Additional Cleaning: all outcome data (dependent variables) will having "unknown" replaced with NA
df[dep_vars] <- lapply(df[dep_vars], gsub, pattern = "Unknown", replacement = NA)

###Converting Character Columns into factors prior to analysis
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], as.factor)

# ALL USED VARIABLES
vars <- c(ind_vars, dep_vars)

#Adjusting Levels in Factors
df$age_group <- relevel(df$age_group, "10 - 19 Years")
df$race_ethnicity_combined <- relevel(df$race_ethnicity_combined, "White, Non-Hispanic")
df$res_region <- relevel(df$res_region, "Northeast")
```
  
* Independent Variables include: `r ind_vars`.  
* Dependent Variables include: `r dep_vars`.  
  
## Data Exploration of Independent and Dependent Variables  
```{r DATA EXPLORATION}
#independent vars table
print("Summary of Independent Variables")
kableone(print(
  CreateTableOne(vars = ind_vars, data = df), 
  showAllLevels = TRUE,
  missing = TRUE,
  noSpaces = TRUE
  ))

#dependent vars
print("Summary of Dependent Variables")
kableone(print(
  CreateTableOne(vars = dep_vars, data = df), 
  showAllLevels = TRUE,
  missing = TRUE,
  noSpaces = TRUE
  ))
```

## Checking Data Variables Prior to Analysis  
```{r CHECK DATA}
#Check data
ff_glimpse(df)
```
  
# Data Analysis  

## Univariate Analysis of Special Populations  
```{r UNIVARIATE-SPECIAL POPULATIONS, results = 'asis'}
## Health Care Workers
print("Univariate Analysis of Health Care Workers")

stratify <- "hc_work_yn"
fischer_var = c()
cat_vars <- colnames(df[,sapply(df, function(x) is.factor(x))])
cat_vars <- cat_vars[cat_vars %in% vars]
for (i in 1:length(cat_vars)) {
  if(sum(chisq.test(table(df[,cat_vars[i]],df[,stratify]))$expected < 5) > 0) {
    fischer_var[i] <- cat_vars[i]
  }
}
fischer_var <- fischer_var[!is.na(fischer_var)]

kableone(print(
  CreateTableOne(vars = vars, data = df, strata = stratify), 
  exact = fischer_var,
  showAllLevels = TRUE,
  missing = TRUE
  ))
cat('\n\n<!-- -->\n\n')



## Previous Medical Conditions
print("Univariate Analysis of Patients with Previous Medical Conditions")

stratify <- "medcond_yn"
fischer_var = c()
cat_vars <- colnames(df[,sapply(df, function(x) is.factor(x))])
cat_vars <- cat_vars[cat_vars %in% vars]
for (i in 1:length(cat_vars)) {
  if(sum(chisq.test(table(df[,cat_vars[i]],df[,stratify]))$expected < 5) > 0) {
    fischer_var[i] <- cat_vars[i]
  }
}
fischer_var <- fischer_var[!is.na(fischer_var)]

kableone(print(
  CreateTableOne(vars = vars, data = df, strata = stratify), 
  exact = fischer_var,
  showAllLevels = TRUE,
  missing = TRUE
  ))
cat('\n\n<!-- -->\n\n')


## Age
print("Univariate Analysis of Age")

stratify <- "age_group"
fischer_var = c()
cat_vars <- colnames(df[,sapply(df, function(x) is.factor(x))])
cat_vars <- cat_vars[cat_vars %in% vars]
for (i in 1:length(cat_vars)) {
  if(sum(chisq.test(table(df[,cat_vars[i]],df[,stratify]))$expected < 5) > 0) {
    fischer_var[i] <- cat_vars[i]
  }
}
fischer_var <- fischer_var[!is.na(fischer_var)]

kableone(print(
  CreateTableOne(vars = vars, data = df, strata = stratify, argsExact = list(workspace = 2*10^5,simulate.p.value=TRUE)), 
  exact = fischer_var,
  showAllLevels = TRUE,
  missing = TRUE
  ))
cat('\n\n<!-- -->\n\n')

## Graph for Age Population
age_graph <- df %>% 
  mutate(age_group = relevel(age_group,"0 - 9 Years")) %>%
  filter(!is.na(df$age_group) & df$age_group != "Unknown") %>%
  group_by(age_group) %>% 
  summarize(
    hosp_yn  = mean(hosp_yn == "Yes", na.rm = TRUE),
    icu_yn  = mean(icu_yn == "Yes", na.rm = TRUE),
    mechvent_yn = mean(mechvent_yn == "Yes", na.rm = TRUE),
    pna_yn = mean(pna_yn == "Yes", na.rm = TRUE),
    acuterespdistress_yn = mean(acuterespdistress_yn == "Yes", na.rm = TRUE),
    death_yn = mean(death_yn == "Yes", na.rm = TRUE)) %>%
  as.data.frame() %>%
  reshape::melt(id =c("age_group")) %>%
  mutate(value = value*100)

ggplot(data=age_graph, aes(x=age_group, y=value, group = variable, color=variable)) +
  geom_point() +
  geom_line() +
  labs(x = "Age Groups", y = "Percent (%)") +
  scale_color_discrete(name = "Outcomes", 
                      breaks = c("hosp_yn","icu_yn","mechvent_yn","pna_yn","acuterespdistress_yn","death_yn"),
                      labels = c("Hospitalization","ICU Admission","Mechanical Ventilation","Pneumonia","ARDS","Death"))

ggsave("figure_age.tiff", width = 9, height = 6.5, units = c("in"), dpi = 1200)


```

## Univariate Analysis by Dependent Variables  
  
* Dependent Variables include: `r dep_vars`.  
  
```{r UNIVARIATE ANALYSIS DEPENDENT, results='asis'}
for (i in 1:length(dep_vars)){
  stratify <- dep_vars[i]
  print(paste("Outcome:", dep_vars[i]))
  
  fischer_var = c()
  cat_vars <- colnames(df[,sapply(df, function(x) is.factor(x))])
  cat_vars <- cat_vars[cat_vars %in% vars]
  for (i in 1:length(cat_vars)) {
    if(sum(chisq.test(table(df[,cat_vars[i]],df[,stratify]))$expected < 5) > 0) {
      fischer_var[i] <- cat_vars[i]
    }
  }
  fischer_var <- fischer_var[!is.na(fischer_var)]
  
cat('\n\n<!-- -->\n\n')
  print(kableone(print(
    CreateTableOne(vars = vars, data = df, strata = stratify, argsExact = list(workspace = 2*10^5,simulate.p.value=TRUE)), 
    exact = fischer_var,
    showAllLevels = TRUE,
    missing = TRUE
  )))
cat('\n\n<!-- -->\n\n')
} 
```
  
## Adjusted Model by Dependent Variables  
  
```{r ADJUSTED MODEL DEPENDENT, results='asis'}
#Analysis

figure.forest = list()

for(i in 1:length(dep_vars)){
#creating reference table
ref.table <- data.frame(
  or = 1,
  x2_5_percent = 1,
  x97_5_percent = 1,
  p = NA,
  variable = c("sexMale", "age_group 10 - 19 Years", "race_ethnicity_combinedWhite, Non-Hispanic", "res_regionNortheast", "hc_work_ynNo", "medcond_ynNo"),
  dependent = dep_vars[i],
  sig_text = NA)
  
  
  print(paste("Outcome:", dep_vars[i]))
  form <- as.formula(paste(dep_vars[i] ,"~", paste0(ind_vars, collapse = " + ")))
  mod <- glm(formula = form, data = df, family = binomial) %>%
    MASS::stepAIC(trace = TRUE, direction = "both")
  
cat('\n\n<!-- -->\n\n')
  print(kableone(print(cbind(exp(cbind(OR = coef(mod), confint.default(mod))), P=summary(mod)$coefficients[,4]))))
cat('\n\n<!-- -->\n\n')

## CREATING OR PLOT
df %>%  
  mutate(race_ethnicity_combined = gsub(", Non-Hispanic", "", race_ethnicity_combined)) %>%
  mutate(res_region = gsub("US Territory or Unknown", "Territories of the United States or Unknown", res_region)) %>%
  mutate(hosp_yn = ff_label(hosp_yn, "Hospitalization")) %>% 
  mutate(icu_yn = ff_label(icu_yn, "ICU Admission")) %>% 
  mutate(mechvent_yn = ff_label(mechvent_yn, "Mechanical Ventilation/Intubation")) %>% 
  mutate(pna_yn = ff_label(pna_yn, "Pneumonia")) %>% 
  mutate(acuterespdistress_yn = ff_label(acuterespdistress_yn, "Acute Respiratory Distress Syndrome")) %>% 
  mutate(death_yn = ff_label(death_yn, "Death")) %>% 
  mutate(sex = ff_label(sex, "Sex")) %>% 
  mutate(age_group = ff_label(age_group, "Age Group")) %>% 
  mutate(race_ethnicity_combined = ff_label(race_ethnicity_combined, "Race & Ethnicity")) %>% 
  mutate(res_region = ff_label(res_region, "United States Census Region")) %>% 
  mutate(hc_work_yn = ff_label(hc_work_yn, "Health Care Worker")) %>% 
  mutate(medcond_yn = ff_label(medcond_yn, "Pre-existing Medical Conditions")) %>% 
  or_plot(dep_vars[i], ind_vars, 
  breaks = c(0.01,0.1, 1, 10, 100, 1000))


figure.forest[[i]] <- cbind(exp(cbind(OR = coef(mod), confint.default(mod))),P=summary(mod)$coefficients[,4]) %>%
  as.data.frame() %>%
  clean_names() %>%
  mutate(variable=rownames(.)) %>%
  `rownames<-`(seq_len(nrow(.))) %>%
  mutate(dependent=dep_vars[i]) %>%
  mutate(sig_text = paste(round(or,2), "(", round(x2_5_percent,2), "-", round(x97_5_percent,2), "), p = ", round(p,2), sep = "")) %>%
  rbind(.,ref.table) %>%
  filter(variable!="(Intercept)") %>%
  arrange(variable) %>%
  add_row(variable= "Sex", dependent=dep_vars[i]) %>%
  add_row(variable= "Age Group", dependent=dep_vars[i]) %>%
  add_row(variable= "Race & Ethnicity", dependent=dep_vars[i]) %>%
  add_row(variable= "Region", dependent=dep_vars[i]) %>%
  add_row(variable= "Healthcare Worker", dependent=dep_vars[i]) %>%
  add_row(variable= "Pre-existing Medical Conditions", dependent=dep_vars[i]) 
  
}




forest <- do.call(dplyr::bind_rows, figure.forest)

ggplot(data=forest, aes(y=variable, x=or, xmin=x2_5_percent, xmax=x97_5_percent)) +
  geom_point() +
  geom_errorbarh(height=.1)+
  scale_x_continuous(limits=c(-0,2), breaks = c(0, 0.5, 1, 1.5, 2), name="OR", oob=scales::squish)  + 
  scale_y_discrete(limits = rev(levels(forest$variable))) +
  geom_vline(xintercept=1, color="black", linetype="dashed", alpha=.5) +
  facet_grid(cols = vars(dependent)) +
  theme(
    axis.ticks.y=element_blank(),
    axis.line.x = element_line(color="black", size = 0.1),
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    plot.background=element_blank()
    )

ggsave("forest_figure.tiff", width = 11, height = 6.5, units = c("in"), dpi = 1200)
```



